---
title: "
<p>Prediction Assignment Writeup"
date: "06.09.16"
output:
  html_document:
    theme: cerulean
---
### Introduction:

In this study I am using data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.The goal of this project is to predict the manner in which they did the exercise. This is the classe variable in the training set. I can use any of the other variables to predict with. I will use my prediction model to predict 20 different test cases.

### Data:

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

### My report:
```{r  include=FALSE}
library(caret)
library(AppliedPredictiveModeling)
```
Reading data
```{r}
training<-read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"), stringsAsFactors = FALSE)
finalTest<-read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"), stringsAsFactors = FALSE)
tsnr<-nrow(finalTest)
trnr<-nrow(training)
finalTest$classe<-1
training$problem_id<-1
```
I am going to bind two datasets to remove Nas and remain only numeric columns
```{r}
trainingSet<-rbind(training,finalTest)
 ```
removing user name
```{r}
trainingSet<-trainingSet[,-1]
```
I am removing columns that are all Nas
```{r}
 trainingSet<-subset(trainingSet, select = colSums(!is.na(trainingSet))>0)
```
In this study I am going to concentrate only on numerical variables
```{r}
training2<-trainingSet[,-112]
nonNumeric=c()
for (name in colnames(training2))
{
    if(is.factor(training2[,name]))     training2[,name]<-as.numeric(as.character(training2[,name]))
    nonNumeric <- c(nonNumeric,is.numeric(training2[,name]))
}
training2<-training2[,nonNumeric]
```
I am removing columns that are all Nas
```{r}
training2<-subset(training2, select = colSums(!is.na(training2))>0)
```
impute missing values with median for test and training set
```{r}
 for (name in names(training2)) {
  training2[,name][is.na(training2[,name])] <-median(as.numeric(training2[,name]), na.rm = TRUE)
 training2[,name] <- ifelse(training2[,name] == 0, median(training2[,name], na.rm = TRUE),                                   training2[,name]) 
}

metric <- "Accuracy"
```
In this case study I will use 10-fold cross validation with 3 repeats
```{r}
 control <- trainControl(method="repeatedcv", number=10, repeats=3)
```
I assign a random number seed to a variable, so that we can re-set the random number generator before I train each algorithm
```{r}
 seed <- 1721
```
adding some noise to data to avoid constant values in columns
```{r}
training2 <- training2 + rnorm(17*3)
```
seperating 20 test cases
```{r}
finalTest<-training2[-(1:trnr),]
training2 <- training2[1:trnr,]
```
divide the training data into test and training
```{r}
inTrain = createDataPartition(training$classe, p = 3/4)[[1]]
pca.train = training2[ inTrain,]
pca.test = training2[-inTrain,]
```
principal component analysis
```{r}
pca.train <- data.frame(t(na.omit(t(pca.train))))
prin_comp <- prcomp(pca.train, scale. = T)
```
compute standard deviation of each principal component
```{r}
std_dev <- prin_comp$sdev
```
compute variance
```{r}
pr_var <- std_dev^2
```
proportion of variance explained
```{r}
prop_varex <- pr_var/sum(pr_var)
plot(prop_varex, xlab = "Principal Component",
             ylab = "Proportion of Variance Explained",
             type = "b")
```
cumulative plot
```{r}
plot(cumsum(prop_varex), xlab = "Principal Component",
       ylab = "Cumulative Proportion of Variance Explained",
       type = "b")
```

This plot shows that 60 components results in variance close to ~ 98%. Therefore, in this case, I will select number of components as 60 [PC1 to PC60] and proceed to the modeling stage.
```{r}
datasetTrain=data.frame(classe = trainingSet[inTrain,"classe"], prin_comp$x)
datasetTrain<-datasetTrain[,1:61]

datasetTrain$classe<-factor(datasetTrain$classe)
```
The algorithms that i was testing are: Linear Discriminant Analysis, Bagged CART, Random Forest, Stochastic Gradient Boosting, Logistic Regression and k-nearest neighbors
kNN gave me the best results.
 
kNN model:
```{r}
set.seed(seed)
fit.knn <- train(classe~., data=datasetTrain, method="knn", metric=metric,  trControl=control)
```
transform test into PCA
```{r}
test.data <- predict(prin_comp, newdata = pca.test)
test.data <- as.data.frame(test.data)
```
select the first 60 components
```{r}
test.data <- test.data[,1:60]
```
make prediction on my test data
```{r}
prediction <- predict(fit.knn, test.data)
cm<-confusionMatrix(prediction, training[-inTrain,"classe"])
cm$overall
```
out of sample error
```{r}
outOfSampleError.accuracy <- sum(prediction == training[-inTrain,"classe"])/length(prediction)
outOfSampleError.accuracy 
```
I will use my prediction model to predict 20 test cases:
```{r}
test.data <- predict(prin_comp, newdata = finalTest)
test.data <- as.data.frame(test.data)
```
select the first 60 components
```{r}
test.data <- test.data[,1:60]
```
make prediction on my test data
```{r}
prediction <- predict(fit.knn, test.data)
prediction
```


### Conclusion:

In my study I compared the accuracy of several algorithms and choose the one that gave me the highest accuracy which was k-nearest neighbors algorithm (k-NN). I choose in my model 60 predictors using PCA and made predictions with this model and 20 provided cases.


